### 2021-09-14 thesis/tex/DCtheory.tex
update

\section{PINGU}
The methodology behind the PINGU simulations are the same as with our DeepCore study~. We use the public MC~\cite{PINGUdata}, which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modelled for PINGU, the correction factors $C_{ijk}$ are all unity.
As with the DeepCore data, the PINGU Monte Carlo is divided into eight 
$\log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. 
We generate `data' by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase which is set to zero for simplicity.

>>>>

\section{PINGU}
The methodology behind the PINGU simulations is the same as with our DeepCore study~. We use the public MC~\cite{PINGUdata}, which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modeled for PINGU, the correction factors $C_{ijk}$ are all unity.
As with the DeepCore data, the PINGU Monte Carlo is divided into eight 
$\log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. 
We generate `data' by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase which is set to zero for simplicity.

### 2021-09-14 thesis/tex/ICintro.tex
update

The strings and DOMs are not spaced evenly, making some parts of the detector more sensitive to certain energy ranges than others.
8 strings packed more tightly than the other 78, making that part of the detector sensitive to neutrino energies down to single-digit \si{\GeV}. Due to 
this part being situated deep within the ice, it is referred to DeepCore. DeepCore will be treated as a separate and independent detector from the rest, which
retains the name IceCube. A view of the current setup can be seen in Fig.~\ref{fig:array}. In this work, we consider DeepCore data between \SIrange{5.6}{56}{\GeV} and IceCube data in the range \SIrange{0.5}{10}{\TeV}.
\begin{figure}\label{fig:array}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/icecube2.png}
    \caption{View of the full IceCube array, with the Eiffel tower for scale.}
\end{figure}

In 2017, the PINGU Letter of Intent was published~\cite{PINGUletter}. The `Precision IceCube Next Generation Upgrade' is an upgrade that will 
supplement DeepCore, i.e. boosing the capabilities of neutrino detection at the \si{\GeV} scale. As the PINGU upgrade is not yet financed nor built, we are
not able to use any data from it. However, the collaboration has released preliminary simulations which we will use to see how the upgrade might improve
results from IceCube and DeepCore. The PINGU simulations have the same structure as the DeepCore data, so our analysis referring to DeepCore will
also apply to PINGU except where noted. However, we treat the PINGU detector as independent of the other experiments.

@@

\begin{table}[h]\label{table:flux}
    \begin{center}
        \begin{tabular}{lcccccc}
            \hline \hline
            $E$ [\si{\GeV}] &$\phi_\mu$ &$\phi_{\bar{\mu}}$ &$\phi_e$ &$\phi_{\bar{e}}$ & $\cos(\theta_z)$\\\\
            \hline
            27825 &  \SI{6.06e-12}{} &  \SI{3.17e-12}{} &  \SI{1.56e-13}{} &  \SI{1.04e-13}{} &   [-0.2, -0.1] \\
            247707 &  \SI{5.94e-16}{} &  \SI{2.92e-16}{} &  \SI{1.36e-17}{} &  \SI{8.12e-18}{} &   [-0.7, -0.6] \\
                22 &  \SI{3.33e-02}{} &  \SI{2.78e-02}{} &  \SI{9.57e-03}{} & \SI{7.15e-03}{} &   [-0.3, -0.2] \\
            432876 &  \SI{5.19e-17}{} &  \SI{2.32e-17}{} &  \SI{1.46e-18}{} & \SI{9.83e-19}{} &   [-1.1, -1.0] \\
            64280 &  \SI{1.58e-13}{} &  \SI{8.10e-14}{} &  \SI{3.49e-15}{} &  \SI{2.21e-15}{} &   [-0.4, -0.3] \\
            \hline \hline
        \end{tabular}
    \end{center}
    \caption{A selection of processed atmospheric South Pole fluxes from~\cite{hondaData} by Honda et al.~\cite{hondaArticle}.}
\end{table}

\begin{figure}[t]\label{fig:flux}
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flux.pdf}
    \caption{Interpolated South Pole atmospheric $\nu_\mu$ flux with data from~\cite{hondaArticle}, displaying a near-neglible 
    zenith angle dependance, but a strong energy dependence. The energy dependence can be parametrized to be approximated by $E^{-2.7}$.}
\end{figure}

>>>>

The strings and DOMs are not spaced evenly, making some parts of the detector more sensitive to certain energy ranges than others.
8 strings packed more tightly than the other 78, making that part of the detector sensitive to neutrino energies down to single-digit \si{\GeV}. Due to 
this part being situated deep within the ice, it is referred to DeepCore. DeepCore will be treated as a separate and independent detector from the rest, which
retains the name IceCube. A view of the current setup can be seen in Fig.~\ref{fig:array}. In this work, we consider DeepCore data between \SIrange{5.6}{56}{\GeV} and IceCube data in the range \SIrange{0.5}{10}{\TeV}.
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/icecube2.png}
    \caption{View of the full IceCube array, with the Eiffel tower for scale.}\label{fig:array}
\end{figure}

In 2017, the PINGU Letter of Intent was published~\cite{PINGUletter}. The `Precision IceCube Next Generation Upgrade' is an upgrade that will 
supplement DeepCore, i.e. boosting the capabilities of neutrino detection at the \si{\GeV} scale. As the PINGU upgrade is not yet financed nor built, we are
not able to use any data from it. However, the collaboration has released preliminary simulations which we will use to see how the upgrade might improve
results from IceCube and DeepCore. The PINGU simulations have the same structure as the DeepCore data, so our analysis referring to DeepCore will
also apply to PINGU except where noted. However, we treat the PINGU detector as independent of the other experiments.

@@

\begin{table}[h]
    \begin{center}
        \begin{tabular}{lcccccc}
            \hline \hline
            $E$ [\si{\GeV}] &$\phi_\mu$ &$\phi_{\bar{\mu}}$ &$\phi_e$ &$\phi_{\bar{e}}$ & $\cos(\theta_z)$\\\\
            \hline
            27825 &  \SI{6.06e-12}{} &  \SI{3.17e-12}{} &  \SI{1.56e-13}{} &  \SI{1.04e-13}{} &   [-0.2, -0.1] \\
            247707 &  \SI{5.94e-16}{} &  \SI{2.92e-16}{} &  \SI{1.36e-17}{} &  \SI{8.12e-18}{} &   [-0.7, -0.6] \\
                22 &  \SI{3.33e-02}{} &  \SI{2.78e-02}{} &  \SI{9.57e-03}{} & \SI{7.15e-03}{} &   [-0.3, -0.2] \\
            432876 &  \SI{5.19e-17}{} &  \SI{2.32e-17}{} &  \SI{1.46e-18}{} & \SI{9.83e-19}{} &   [-1.1, -1.0] \\
            64280 &  \SI{1.58e-13}{} &  \SI{8.10e-14}{} &  \SI{3.49e-15}{} &  \SI{2.21e-15}{} &   [-0.4, -0.3] \\
            \hline \hline
        \end{tabular}
    \end{center}
    \caption{A selection of processed atmospheric South Pole fluxes from~\cite{hondaData} by Honda et al.~\cite{hondaArticle}.}\label{table:flux}
\end{table}

\begin{figure}[t]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flux.pdf}
    \caption{Interpolated South Pole atmospheric $\nu_\mu$ flux with data from~\cite{hondaArticle}, displaying a near-neglible 
    zenith angle dependance, but a strong energy dependence. The energy dependence can be parametrized to be approximated by $E^{-2.7}$.}\label{fig:flux}
\end{figure}

### 2021-09-14 thesis/tex/nsiTheory.tex
update

% \begin{document}
\section{Non-Standard Interactions}\label{sec:nsiTheory}
We have no reason to believe that the Standard Model 
gauge group is the complete picture. We have already seen how the failure of the Standard Model to predict neutrino masses and flavor 
osillations forces us to amend it. 
Just as the electroweak theory $SU(2)_L \times U(1)_Y$ is 
spontaneously broken to $U(1)_{EM}$, a higher order theory at a different energy scale with completely different
properties might undergo a similar spontaneous symmetry break at high energies, 
producing at lower energies the Standard Model that we know. In fact, 
just as the fermion masses originate from the electroweak symmetry breaking, the neutrino masses might be generated from 
another broken symmetry resulting in the Standard Model gauge group.
In this sense, the Standard Model might be considered 
an effective low energy theory, which is yiels impressive results in some areas, but failing in others.

@@

\subsection{IceCube Signal}
In our analysis of IceCube, we are constrained to muon track events. Thus, we are not able to test any theory which does not modify $P_{\a \mu}$. Moreover,
the IceCube data is avaibable in the range \SI{500}{\GeV} to \SI{10}{\TeV} range, where any rapid oscillations have settled down.

Since all standard matter potentials are diagonal, 
the elements $\epsilon_{\a\b},\, \a = \b$ will directly adjust the matter potential felt 
by flavor $\a$. The off-diagonal terms have a more interesting theoretical implication as they open up matter interactions across flavors. Remember, in the Standard model,
we are restriced to weak interactions that conserve lepton number. However, a off-diagonal NSI parameter allows flavor transitions during matter interactions. Thus,
the off-diagonal elements constitute new soures of flavor violation.
Moreover, $\epsilon_{\a\b}$ modifies the $\nu_\a \to \nu_\b$ transition independently
of the mixing matrix. Thus, we can in theory boost a flavor transition which is supressed by the mixing matrix. Moreover, since we have 6 NSI parameters, and the PMNS
matrix is parametrized with only 3 parameters, we have more degrees of freedom when adjusting the probabilities using the NSI matrix. Each term in the PMNS matrix consists of at least
two of the three mixing angles, making them much more dependent on each other.

As discussed in Eq.~\ref{eq:Uvalues}, the atmospheric $\nm \to \nt$ transition will be the most abundant, making $\emt$, $\emm$, $\ett$ the most suitable
NSI parameters to constrain from muon events. As we will see, $\eem$ is also a candidate, alebit a weaker one.

In Fig.~\ref{fig:emt_ett_probs}, we see how the introduction of $\emt = 0.02$ alters the $\nm$ and $\anm$ survival probabilities
for neutrinos that traverse the entire Earth diameter (i.e. $\ztrue = -1$). $\emt$ does not dramatically change neither
amplitude nor frequency of the probabilities. Instead, it seems to strech or compress the oscillations. Since the 
only difference between the way neutrinos and antineutrinos interact with matter is the sign of the potential, the probability for
$\nm$ with positive $\epsilon_{\a\b}$ is identical to the probability for $\anm$ with negative $\epsilon_{\a\b}$. Thus, the dashed line 
in the right panel not only shows the survival probability for $\anm$ with $\emt=0.02$, but also the survival probability for $\nm$ with $\emt=-0.02$.
Hence, we note that $\emt > 0$ stretches (compresses) $\Pmm$ for neutrinos (antineutrinos), while $\emt < 0$ compresses (stretches) $\Pmm$ for neutrinos (antineutrinos).

The value of $\ett$ does not affect neither $\Pmm$ nor $\Pamam$, in the IceCube region above \SI{500}{\GeV}. Hence, we will not be able
to say anything about $\ett$ in our IceCube study.  Comparing the probabilities in Fig.~\ref{fig:emt_ett_probs} with $\ett = 0.05$ with the ones for $\emt = 0.02$ in Fig.~\ref{fig:emt_ett_probs},
we see that even though we let $\ett$ take 2.5 times the value of $\emt$, its effect on $\Pmm$ is smaller. The weakening of the $\Pamam$ resonance will be visible in DeepCore, but we should expect a less stringent 
constraint due to the weakness of the effect compared to $\emt$.

@@

Moving on to $\eem$ and Fig.~\ref{fig:eem_eet_probs}, we see that both probabilities has shifted downwards for $\Etrue > \SI{500}{\GeV}$.
In Fig.~\ref{fig:eem_eet_probs}, we see that the muon channel remains largely unaffected of the value of $\eet$ as we expected. The expection of this lies 
in the DeepCore region of rapid oscillations, where mixing is more violent. 
\begin{figure}
    \begin{center}%TODO subplots and fix labels in parahgraph aboveto match them
        \includegraphics[width=0.49\textwidth]{figures/emt_probs.pdf}
        \includegraphics[width=0.49\textwidth]{figures/ett_probs.pdf}
        \caption{\emph{Left panel:} Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\emt = 0.02$. All other NSI parameters are fixed to zero.
        \emph{Right panel:} Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\ett = 0.05$. All other NSI parameters are fixed to zero.}
        \label{fig:emt_ett_probs}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
        \includegraphics[width=0.49\textwidth]{figures/eem_probs.pdf}
        \includegraphics[width=0.49\textwidth]{figures/eet_probs.pdf}
        \caption{Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$\emph{Left panel:} $\eem = 0.2$. All other NSI parameters are fixed to zero.
        \emph{Right panel:} $\eet = 0.2$. All other NSI parameters are fixed to zero.}
        \label{fig:eem_eet_probs}
    \end{center}
\end{figure}

\subsection{DeepCore and PINGU Signals}
Now we repeat our proability analysis but for the DeepCore/PINGU region of \SIrange{5.6}{56}{\GeV}. As we previously saw,
we have rapid oscillations, which means that `indirect' modifications (i.e. $\eet$ will affect the $\Pmm$ channel)
will be more apparent, since all flavors are involved to a greater degree compared with the more stable region above \SI{500}{\GeV}, where 
many oscillations have settled down.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.
%TODO: clean up above and expand a bit.
\begin{figure}
    \begin{center}
        \includegraphics[width=0.8\textwidth]{figures/Pee_eet_probs.pdf}
        \caption{Electron neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\eet = 0.2$. All other NSI parameters are fixed to zero.}
        \label{fig:Pee_eet_probs}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
        \includegraphics[width=0.8\textwidth]{figures/DC_event_pulls.pdf}
        \caption{Plots of the quantity $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$, displaying which 
        reconstructed bins the NSI signal is most clear.}
        \label{fig:event_pulls}
    \end{center}
\end{figure}
%TODO: talk a bit about this. Compare cascades to tracks

>>>>

% \begin{document}
\section{Non-Standard Interactions}\label{sec:nsiTheory}
We have no reason to believe that the Standard Model 
gauge group is the complete picture. We have already seen how the failure of the Standard Model to predict neutrino masses and flavor 
oscillations force us to amend it. 
Just as the electroweak theory $SU(2)_L \times U(1)_Y$ is 
spontaneously broken to $U(1)_{EM}$, a higher order theory at a different energy scale with completely different
properties might undergo a similar spontaneous symmetry break at high energies, 
producing at lower energies the Standard Model that we know. In fact, 
just as the fermion masses originate from the electroweak symmetry breaking, the neutrino masses might be generated from 
another broken symmetry resulting in the Standard Model gauge group.
In this sense, the Standard Model might be considered 
an effective low energy theory, which is yields impressive results in some areas, but failing in others.

@@

\subsection{IceCube Signal}
In our analysis of IceCube, we are constrained to muon track events. Thus, we are not able to test any theory which does not modify $P_{\a \mu}$. Moreover,
the IceCube data is available in the range \SI{500}{\GeV} to \SI{10}{\TeV} range, where any rapid oscillations have settled down.

Since all standard matter potentials are diagonal, 
the elements $\epsilon_{\a\b},\, \a = \b$ will directly adjust the matter potential felt 
by flavor $\a$. The off-diagonal terms have a more interesting theoretical implication as they open up matter interactions across flavors. Remember, in the Standard model,
we are restricted to weak interactions that conserve lepton number. However, an off-diagonal NSI parameter allows flavor transitions during matter interactions. Thus,
the off-diagonal elements constitute new sources of flavor violation.
Moreover, $\epsilon_{\a\b}$ modifies the $\nu_\a \to \nu_\b$ transition independently
of the mixing matrix. Thus, we can in theory boost a flavor transition that is suppressed by the mixing matrix. Moreover, since we have 6 NSI parameters, and the PMNS
matrix is parametrized with only 3 parameters, we have more degrees of freedom when adjusting the probabilities using the NSI matrix. Each term in the PMNS matrix consists of at least
two of the three mixing angles, making them much more dependent on each other.

As discussed in Eq.~\ref{eq:Uvalues}, the atmospheric $\nm \to \nt$ transition will be the most abundant, making $\emt$, $\emm$, $\ett$ the most suitable
NSI parameters to constrain from muon events. As we will see, $\eem$ is also a candidate, albeit a weaker one.

In Fig.~\ref{fig:emt_ett_probs}, we see how the introduction of $\emt = 0.02$ alters the $\nm$ and $\anm$ survival probabilities
for neutrinos that traverse the entire Earth diameter (i.e. $\ztrue = -1$). $\emt$ does not dramatically change neither
amplitude nor frequency of the probabilities. Instead, it seems to stretch or compress the oscillations. Since the 
only difference between the way neutrinos and antineutrinos interact with matter is the sign of the potential, the probability for
$\nm$ with positive $\epsilon_{\a\b}$ is identical to the probability for $\anm$ with negative $\epsilon_{\a\b}$. Thus, the dashed line 
in the right panel not only shows the survival probability for $\anm$ with $\emt=0.02$, but also the survival probability for $\nm$ with $\emt=-0.02$.
Hence, we note that $\emt > 0$ stretches (compresses) $\Pmm$ for neutrinos (antineutrinos), while $\emt < 0$ compresses (stretches) $\Pmm$ for neutrinos (antineutrinos).

The value of $\ett$ affects neither $\Pmm$ nor $\Pamam$, in the IceCube region above \SI{500}{\GeV}. Hence, we will not be able
to say anything about $\ett$ in our IceCube study.  Comparing the probabilities in Fig.~\ref{fig:emt_ett_probs} with $\ett = 0.05$ with the ones for $\emt = 0.02$ in Fig.~\ref{fig:emt_ett_probs},
we see that even though we let $\ett$ take 2.5 times the value of $\emt$, its effect on $\Pmm$ is smaller. The weakening of the $\Pamam$ resonance will be visible in DeepCore, but we should expect a less stringent 
constraint due to the weakness of the effect compared to $\emt$.

@@

Moving on to $\eem$ and Fig.~\ref{fig:eem_eet_probs}, we see that both probabilities has shifted downwards for $\Etrue > \SI{500}{\GeV}$.
In Fig.~\ref{fig:eem_eet_probs}, we see that the muon channel remains largely unaffected of the value of $\eet$ as we expected. The expection of this lies 
in the DeepCore region of rapid oscillations, where mixing is more violent. 
\begin{figure}
    \begin{center}%TODO subplots and fix labels in parahgraph aboveto match them
        \includegraphics[width=0.9\textwidth]{figures/emt_probs.pdf}
        \includegraphics[width=0.9\textwidth]{figures/ett_probs.pdf}
        \caption{\emph{Top panel:} Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\emt = 0.02$. All other NSI parameters are fixed to zero. In the \si{\GeV} range, $\emt$ shifts the oscillations to the right for $\nm$, and to the left for $\anm$. At \si{\TeV} energies, both probabilites simply get shifted down, resulting in a net reduction of track events.
        \emph{Bottom panel:} Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\ett = 0.05$. All other NSI parameters are fixed to zero. $\ett$ does not affect the probabilities above \SI{100}{\GeV} and this parameter is thus unable to be constrained by tracks in IceCube in our study. However, the dampening of the $\anm$ survival probability will be visible to DeepCore and PINGU, since it occurs within their energy ranges.}
        \label{fig:emt_ett_probs}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
        \includegraphics[width=0.9\textwidth]{figures/eem_probs.pdf}
        \includegraphics[width=0.9\textwidth]{figures/eet_probs.pdf}
        \caption{Muon neutrino and antineutrino survival probabilities for
        $\ztrue = -1$. \emph{Top panel:} $\eem = 0.2$. All other NSI parameters are fixed to zero. Instead of the oscillations shifting to the left or right, $\eem$ dampens an oscillation peak for low \si{\GeV}. At \si{\TeV}, the probability is shifted down, just as with $\emt$.
        \emph{Bottom panel:} $\eet = 0.2$. All other NSI parameters are fixed to zero. Here, we see a very weak shifting and a $\nm$ dampening at low \si{\GeV}, and again, no visible effect at \si{TeV} energies. }
        \label{fig:eem_eet_probs}
    \end{center}
\end{figure}

\subsection{DeepCore and PINGU Signals}
Now we repeat our probability analysis but for the DeepCore/PINGU region of \SIrange{5.6}{56}{\GeV}. As we previously saw,
we have rapid oscillations, which means that `indirect' modifications (i.e. $\eet$ will affect the $\Pmm$ channel)
will be more apparent, since all flavors are involved to a greater degree compared with the more stable region above \SI{500}{\GeV}, where 
many oscillations have settled down.

@@

We plot the event pull $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ where $N_{(N)SI}$ are the numbers of expected events
assuming (non-)standard interactions in Fig.~\ref{fig:event_pulls}. This gives the normalized difference in the
number of expected events at the detector, and illustrates the expected sensitivity of DeepCore for the NSI parameters.
%TODO: clean up above and expand a bit.
\begin{figure}
    \begin{center}
        \includegraphics[width=0.9\textwidth]{figures/Pee_eet_probs.pdf}
        \caption{Electron neutrino and antineutrino survival probabilities for
        $\ztrue = -1$ when $\eet = 0.2$. All other NSI parameters are fixed to zero. Compared to the $\nm$ and $\anm$ plots, here we see a strong difference across the whole energy range.}
        \label{fig:Pee_eet_probs}
    \end{center}
\end{figure}

\begin{figure}
    \begin{center}
        \includegraphics[width=0.8\textwidth]{figures/DC_event_pulls.pdf}
        \caption{Plots of the quantity $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$, displaying which 
        reconstructed bins the NSI signal is most clear.}
        \label{fig:event_pulls}
    \end{center}
\end{figure}

Now we will compare the probability plots with the final event counts in DeepCore. It is not enough to only study the effect at probability level, since that is in true quantities. Since the detector data comes in reconstructed quantities, it might be the case that a feature at probability level gets smeared away in the reconstruction, and not showing up at all in the data. In general, features showing at probability level are often washed out at reconstruction if they occur in the rapidly oscillating part of single-digit \si{\GeV} energies.

The top right panel shows a surplus of DeepCore track events at \SIrange{20}{30}{\GeV} for through-going neutrinos when we turn on $\ett$. This is what we expected from the bottom right panel in Fig.~\ref{fig:emt_ett_probs}, where we saw that we had an increase in $\anm$ survival probability at \SIrange{20}{40}{\GeV}.

The third right panel shows the expected event pull for $\eem$. Comparing with the probabilities in Fig.~\ref{fig:eem_eet_probs}, we see that the surplus of the $\anm$ survival in the \SIrange{20}{30}{\GeV} and the deficit above \SI{40}{\GeV} is intact, while the \SI{10}{\GeV} $\nm$ deficit is completely washed out.
